A new tidy data structure to support exploration and modeling of temporal data

### 2019-08-27 tsibble.Rmd
minor tweaks and adjusted gap plots

Time series can be univariate or multivariate, and require relatively long lengths (i.e., large $T$) for modeling. With this large $T$ property, the series can be handled as stochastic processes for the primary purpose of forecasting, and characterizing temporal dynamics. Due to an expectation of regular time intervals, and equal length across series, multivariate time series are typically assumed to be in the format where each column contains a single time series, and time is specified implicitly. This also implies that data are columns of homogeneous types: either all numeric or all non-numeric. To wrestle data, in whatever format that it arrives in, to this modeling format can be frustrating. The format could be considered to be model-centric, rather than data-centric, and thus throws the analyst into the deep end of the pool, rather than allowing them to gently wade to the modeling stage from the shallow end. The expectation is that the "model" is at the center of the analytical universe. This is contrary to the **tidyverse** conceptualization (Figure \@ref(fig:tsibble-fit)), which holistically captures the full data workflow. More support needs to be provided, in the form of consistent tools and data structures, to transform the data into the analytical cycle.

@@

Stata [@Stata] provides two commands, `tsset` and `xtset`, to declare time series and panel data respectively, both of which require explicit panel and time index specification.  The underlying data arrangement is only long form, for both types of data. The functions pay no attention to whether the data is for time series analysis or longitudinal data analysis. The SAS software [@SAS-STAT] also handles both types of data in the same way as Stata.

>>>>

Time series can be univariate or multivariate, and require relatively long lengths (i.e., large $T$) for modeling. With this large $T$ property, the series can be handled as stochastic processes for the primary purpose of forecasting, and characterizing temporal dynamics. Due to an expectation of regularly spaced time, and equal lengths across series, multivariate time series are typically assumed to be in the format where each column contains a single time series, and time is specified implicitly. This also implies that data are columns of homogeneous types: either all numeric or all non-numeric. To wrestle data, in whatever format that it arrives in, to this modeling format can be frustrating. The format could be considered to be model-centric, rather than data-centric, and thus throws the analyst into the deep end of the pool, rather than allowing them to gently wade to the modeling stage from the shallow end. The expectation is that the "model" is at the center of the analytical universe. This is contrary to the **tidyverse** conceptualization (Figure \@ref(fig:tsibble-fit)), which holistically captures the full data workflow. More support needs to be provided, in the form of consistent tools and data structures, to transform the data into the analytical cycle.

@@

Stata [@Stata] provides two commands, `tsset` and `xtset`, to declare time series and panel data respectively, both of which require explicit panel and time index specification. The underlying data arrangement is only long form, for both types of data. Both groups of functions can be applied interchangeably to whether the data is declared for time series or longitudinal data. The SAS software [@SAS-STAT] also handles both types of data in the same way as Stata.
